---
title: "Confirmed Ebola Cases In Three Countries"
author: "Anna-Sophia Stocker"
date: "`r date()`"
format: 
  html:
   toc: true
   code-fold: show
   code-tools: true
   df-print: paged
   embed-resources: true
   
---

```{r}
#| output: false
# Load libraries
library(here)
library(tidyverse)
library(lubridate)
```

## Introduction

The aim of this exercises is to visualize the cumulative number of confirmed ebola cases in three African countries in 2014 and 2015.

## Data

We read and process the data of confirmed ebola cases.

```{r}
#| output: false
# Read data
ebola <- read.csv("C:/Users/Home-Office/Documents/PHS-Kurs/basicstatistics/data/raw/ebola.csv")
```

```{r}
# Process data
ebola$Date <- as.Date(ebola$Date) 
ebola <- arrange(ebola, Date)
ebolacases <- ebola |>
  select(date = Date, country = Country, cum_conf_cases = Cum_conf_cases) |>
  filter(date <= as.Date("2015-03-31") & 
           (country == "Guinea" | country == "Liberia" | country == "Sierra Leone"))
```

## Results

@fig-cases shows the cumulative number of confirmed ebola cases from August 2014 up to and including March 2015.

```{r}
#| label: fig-cases
#| fig-cap: "confirmed ebola cases in three countries"
#| fig-width: 8
#| fig-height: 4
ggplot(data = ebolacases,  
       mapping = aes(x = date, y =cum_conf_cases, fill = country, colour = country))+
  geom_point(alpha = 0.7, shape = 21, size = 2, stroke = 1)+
  geom_line() +
  scale_fill_manual(name = "country",
                    breaks = c("Guinea", "Liberia", "Sierra Leone"), 
                    values = c("green", "blue", "turquoise"),
                    labels = c("Guinea", "Liberia", "Sierra Leone"))+ 
  scale_colour_manual(name = "country",
                      breaks = c("Guinea", "Liberia", "Sierra Leone"),
                      values = c("green", "blue", "turquoise"),
                      labels = c("Guinea", "Liberia", "Sierra Leone"))+
  scale_x_date(breaks = as.Date(c("2014-08-29", "2014-10-01", "2014-12-01", "2015-02-01", "2015-04-01")),
               labels = c("29 August", "1 October", "1 Dezember", "1 February", "1 April"),
               limits = as.Date(c("2014-08-28", "2015-04-02"))) +
   ggtitle(label = "Confirmed Ebola cases") +
  xlab(label = "Time") +
  ylab(label = "Cum. # of confirmed cases")
```

## Conclusions

Sierra Leone had more than double the cumulative number of cases than Liberia or Guinea in March 2015. The increase of cases in all three countries has slowed down in all three countries since January 2015.
