---
title: "Cost of Health Insurance"
author: "Anna-Sophia Stocker"
date: "`r date()`"
format: 
  html:
   toc: true
   code-fold: show
   code-tools: true
   df-print: paged
   embed-resources: true
---

```{r}
#| output: false
# Load libraries
library(here)
library(tidyverse)
library(rstatix)
library(unibeCols)
library(gtsummary)
```

## Introduction

The aim of this study is to compare the health insurance charges from women and men in the US.

## Data

The data used for this analysis is from the Medical Cost Personal Datasets, that are available open source on [Kaggle](https://www.kaggle.com/datasets/mirichoi0218/insurance?resource=download). We concentrate on the Charges-Data, the amount of health costs accumulated, by women and men in the US. Other available variables are ignored at this point. 

```{r}
#| output: false
# Read and transform data
insurance <- read_csv(here("data/raw/insurance_with_date.csv"))
insurance$sex <- as.factor(insurance$sex)
str(insurance)
```

## Results

### Descriptive Statistics

@tbl-one shows the basic descriptive values of the health insurance charges separate by men and women. 

```{r}
#| label: tbl-one
#| tbl-cap: "Overview of the data on health charges for women and men"

insurance |> 
  tbl_summary(
    by = sex, 
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c("{mean} ({sd})", "{median}", "{p25} - {p75}", "{min}", "{max}"),
    include = c("sex", "charges")
  )
```

A Boxplot (@fig-one) shows the distribution of charges by women and men. On first glance, the data for women and men look quite similar, but with the distribution skewed to the lower end. 

```{r}
#| label: fig-one
#| fig-cap: "Boxplot of the distribution of health charges by sex"
#| fig-subcap: "Whiskers at 1.5 IQR"
ggplot(data = insurance, aes (x = sex, y = charges, fill = sex))+
  geom_boxplot(alpha = 0.7) +
    scale_fill_manual(name = "sex",
                  breaks = c("male", "female"),
                  values = c(unibeApricotS()[1], unibeMintS()[1]),
                  labels = c("male", "female")) +
    scale_colour_manual(name = "sex",
                  breaks = c("male", "female"),
                  values = c(unibeApricotS()[1], unibeMintS()[1]),
                  labels = c("male", "female"))+
  ylab(label = "Charges ($)") +
  xlab("Sex") +
  theme_minimal() + 
    theme(legend.position = "none")
  
```

#### Analysis of normality

The Histogram (@fig-two) also shows a skewed to the left distribution of the Charges-Data in both women and men. The QQ-Plot (@fig-three) indicates that the the distribution differs from the normal distribution. This is confirmed by a Shapiro-Test for Normality (@tbl-two).

```{r}
#| label: fig-two
#| fig-cap: "Histogram of the distribution of health charges by sex"
# Histogram data by sex
ggplot(data = insurance, aes(x = charges, fill = sex)) +
  geom_histogram(alpha = 0.7, binwidth = 1000)+
  facet_wrap("sex")+
  scale_fill_manual(name = "sex",
                    breaks = c("male", "female"),
                    values = c(unibeApricotS()[1], unibeMintS()[1]),
                    labels = c("male", "female")) +
  scale_colour_manual(name = "sex",
                      breaks = c("male", "female"),
                      values = c(unibeApricotS()[1], unibeMintS()[1]),
                      labels = c("male", "female"))+
  theme_minimal() + 
  theme(legend.position = "none")
```

```{r}
#| label: fig-three
#| fig-cap: "QQ-Plot of health-charges data"
# QQ-Plot data by sex

insurance |>
  ggplot(aes(sample = charges, colour = sex, fill = sex)) +
  geom_qq_line(distribution = stats::qnorm)+
  geom_qq(alpha = 0.7)+
  facet_wrap("sex")+
  scale_fill_manual(name = "sex",
                  breaks = c("male", "female"),
                  values = c(unibeApricotS()[1], unibeMintS()[1]),
                  labels = c("male", "female")) +
  scale_colour_manual(name = "sex",
                      breaks = c("male", "female"),
                      values = c(unibeApricotS()[1], unibeMintS()[1]),
                      labels = c("male", "female"))+
  ylab(label = "Sample quantiles") +
  xlab("Theoretical quantiles") +
  theme_minimal() + 
  theme(legend.position = "none")
```

```{r}
#| label: tbl-two
#| tbl-cap: "Shapiro-test for Normality"
# Test for Normality data by sex
insurance |>
  group_by(sex) |>
  shapiro_test(charges) 
```

#### Transformation of data
A log-transformation of the Charges-Data was performed and Histogram (@fig-four) and QQ-Plot (@fig-five) repeated with the transformed Data. The distribution now looks closer to a normal distribution, although a Shapiro-Test for Normality (@tbl-three) still indicates non-normality.

```{r}
#| output: false
# Transform data
insurance$logcharges <- log(insurance$charges)
```

```{r}
#| label: fig-four
#| fig-cap: "Histogram of the distribution of the log-transformed health charges by sex"
# Histogram of transformed data
ggplot(data = insurance, aes(x = logcharges, fill = sex)) +
  geom_histogram(alpha = 0.7, binwidth = 0.1)+
  facet_wrap("sex")+
  scale_fill_manual(name = "sex",
                    breaks = c("male", "female"),
                    values = c(unibeApricotS()[1], unibeMintS()[1]),
                    labels = c("male", "female")) +
  scale_colour_manual(name = "sex",
                      breaks = c("male", "female"),
                      values = c(unibeApricotS()[1], unibeMintS()[1]),
                      labels = c("male", "female"))+
  theme_minimal() + 
  theme(legend.position = "none")
```

```{r}
#| label: fig-five
#| fig-cap: "QQ-Plot of log-transformed health-charges data"
# QQ-Plot of transformed data
insurance |>
  ggplot(aes(sample = logcharges, ill = sex, colour = sex)) +
  geom_qq_line(distribution = stats::qnorm)+
  geom_qq(alpha = 0.7)+
  facet_wrap("sex")+
  scale_fill_manual(name = "sex",
                  breaks = c("male", "female"),
                  values = c(unibeApricotS()[1], unibeMintS()[1]),
                  labels = c("male", "female")) +
  scale_colour_manual(name = "sex",
                      breaks = c("male", "female"),
                      values = c(unibeApricotS()[1], unibeMintS()[1]),
                      labels = c("male", "female"))+
  theme_minimal() + 
  theme(legend.position = "none")
```

```{r}
#| label: tbl-three
#| tbl-cap: "test for normality on transformed data"
# Test for normality on transformed data
insurance |>
  group_by(sex) |> 
  shapiro_test(logcharges)
```

#### Inferential statistics
A t-Test (@tbl-four) on the transformed data and a Wilcoxon rank sum test (@tbl-five) on the non-transformed data were executed. Both tests are not statistically significant. 


```{r}
#| label: tbl-four
#| tbl-cap: "t-test on transformed health charges data"
# T-test on transformed data 
insurance |>
    t_test(logcharges ~ sex)
```

```{r}
#| label: tbl-five
#| tbl-cap: "wilcoxon rank sum test on original health charges data"
# Test on original data 
insurance |>
  wilcox_test(charges ~ sex)

```

## Conclusions

There is no indication of a statistical connection between the amount of health charges a person accumulates and their sex in the US. 
